Overcoming multiple sources of misspecification\\in GWAS summary statistics

### 2021-05-30 paper/paper-misspec-supp.tex
add investigate mis N

@@

>>>>

\FloatBarrier

@@

\FloatBarrier

\begin{figure}[p]
	\begin{subfigure}{\textwidth}
		\centering
		\includegraphics[width=.8\linewidth]{lassosum2-mis-maxN}
	\end{subfigure}\vspace{2em}
	\begin{subfigure}{\textwidth}
		\centering
		\includegraphics[width=.8\linewidth]{lassosum2-mis-imputeN}
	\end{subfigure}
	\caption{Multiple results (squared correlation r2 between polygenic score and phenotype) over a grid of parameters ($\lambda$ and $\delta$) for lassosum2 in one of the simulations with misspecified GWAS sample sizes (N). The top subfigure is using the maximum N, whereas the bottom subfigure is using the imputed N (cf.\ figure \ref{fig:simu-misN}).  Missing points represent models that were detected as divergent.}
	\label{fig:lassosum2-misN}
\end{figure}

\begin{figure}[p]
	\begin{subfigure}{\textwidth}
		\centering
		\includegraphics[width=.8\linewidth]{ldpred2-mis-maxN}
	\end{subfigure}\vspace{2em}
	\begin{subfigure}{\textwidth}
		\centering
		\includegraphics[width=.8\linewidth]{ldpred2-mis-imputeN}
	\end{subfigure}
	\caption{Multiple results (squared correlation r2 between polygenic score and phenotype) over a grid of parameters (the SNP heritability h2 and the proportion of causal variants p) for LDpred2(-grid) in one of the simulations with misspecified GWAS sample sizes (N). The top subfigure is using the maximum N, whereas the bottom subfigure is using the imputed N (cf.\ figure \ref{fig:simu-misN}). Note that the small h2 values (0.002, 0.02, 0.06) were not considered in the results of figure \ref{fig:simu-misN}. Missing points represent models that were detected as divergent.}
	\label{fig:ldpred2-misN}
\end{figure}

\FloatBarrier

### 2021-05-30 paper/paper-misspec.tex
add investigate mis N

We design simulations where variants have different GWAS sample sizes, which is often the case when meta-analyzing GWAS from multiple cohorts without the same genome coverage.
Using 20,000 variants from chromosome 22 (Methods), we simulate quantitative phenotypes assuming an heritability of 20\% and 2000 causal variants.
For half of the variants, we use 100\% of 300,000 individuals for GWAS, but use only 80\% for one quarter of them and 60\% for the remaining quarter. 
We then run C+T, LDpred2-inf, LDpred2(-grid), LDpred2-auto, lassosum, lassosum2, PRS-CS and SBayesR by using either the true per-variant GWAS sample size, or the total sample size. 
We do not show results for SBayesR because it always diverged and for PRS-CS because the overlap with the LD reference they provide is too small.
Averaged over 10 simulations, when providing true per-variant GWAS sample sizes, squared correlations (in \%) between the polygenic scores and the simulated phenotypes are of [TODO: COMPLETE] (Figure \ref{fig:simu-misN}).
Results for LDpred2-auto are the same as with LDpred2-inf, as it always converges to an infinitesimal model (p = 1) in these simulations.
Note that C+T does not use this sample size information.
For methods that use the per-variant GWAS sample sizes, predictive performance slightly decreases [TODO: REPLACE NUMBERS] from 17.5 to 17.4 for lassosum, from 18.0 to 17.5 for lassosum2, but dramatically decreases for LDpred2 with new values of 11.5 for LDpred2-grid, 5.7 for LDpred2-auto and 6.8 for LDpred2-inf (Figure \ref{fig:simu-misN}).
Therefore, this extreme simulation scenario shows that LDpred2 can be sensitive to GWAS sample size misspecification, whereas lassosum (and lassosum2) seems little affected by this.

[TODO: COMPLETE WITH LOOK AT EACH OF THEM]

@@

Among genetic variants with a minor allele frequency larger than 0.01 and an imputation INFO score larger than 0.3, we sample 20,000 of them according to the inverse of the INFO score density so that they have varying levels of imputation accuracy (Figure \ref{fig:hist-info}).
We read the UK Biobank data into two different datasets using function \texttt{snp\_readBGEN} from R package bigsnpr \cite[]{prive2017efficient}, one by reading the BGEN data at random according to imputation probabilities, and another one reading it as dosages (i.e.\ expected values according to imputation probabilities).
The first dataset is used as what could be the real genotype calls and the second dataset as what would be its imputed version; this technique was used in \cite{prive2019making}.

>>>>

We design simulations where variants have different GWAS sample sizes, which is often the case when meta-analyzing GWAS from multiple cohorts without the same genome coverage.
Using 20,000 variants from chromosome 22 (Methods), we simulate quantitative phenotypes assuming an heritability of 20\% and 2000 causal variants.
For half of the variants, we use 100\% of 300,000 individuals for GWAS, but use only 80\% for one quarter of them and 60\% for the remaining quarter. 
We then run C+T, LDpred2-inf, LDpred2(-grid), LDpred2-auto, lassosum, lassosum2, PRS-CS and SBayesR by using either the true per-variant GWAS sample size, or the total sample size. 
We do not show results for SBayesR because it always diverged and for PRS-CS because the overlap with the LD reference they provide is too small.
Averaged over 10 simulations, when providing true per-variant GWAS sample sizes, squared correlations (in \%) between the polygenic scores and the simulated phenotypes are of [TODO: COMPLETE] (Figure \ref{fig:simu-misN}).
Note that C+T does not use this sample size information.
For methods that use the per-variant GWAS sample sizes, predictive performance slightly decreases [TODO: REPLACE NUMBERS] from 17.5 to 17.4 for lassosum, from 18.0 to 17.5 for lassosum2, but dramatically decreases for LDpred2 with new values of 11.5 for LDpred2-grid, 5.7 for LDpred2-auto and 6.8 for LDpred2-inf (Figure \ref{fig:simu-misN}).
Therefore, this extreme simulation scenario shows that LDpred2 can be sensitive to GWAS sample size misspecification, whereas lassosum (and lassosum2) seems little affected by this.

We then conduct further investigations to explain results of figure \ref{fig:simu-misN}.
First, the reason why results for LDpred2-auto are the same as with LDpred2-inf is because it always converges to an infinitesimal model (p = 1) in these simulations.
Second, for lassosum2, results for a grid of parameters (over $\lambda$ and $\delta$) and quite smooth compared to LDpred2 (Figures \ref{fig:lassosum2-misN} and \ref{fig:ldpred2-misN}). Results slightly improve globally for both lassosum2 and LDpred2 when using imputed instead of maximum GWAS sample sizes.
Finally, in these simulations with misspecified sample sizes, it seems highly beneficial to use a small value for the SNP heritability hyper-parameter in LDpred2, e.g.\ a value of 0.02 or even 0.002 when the true value is 0.2 (Figure \ref{fig:ldpred2-misN}). 
We recall that using a small value for this hyper-parameter induces a larger regularization on the effect sizes, and this benefits LDpred2 even when using imputed sample sizes.

@@

Among genetic variants with a minor allele frequency larger than 0.01 and an imputation INFO score larger than 0.4, we sample 40,000 of them according to the inverse of the INFO score density so that they have varying levels of imputation accuracy (Figure \ref{fig:hist-info}).
We read the UK Biobank data into two different datasets using function \texttt{snp\_readBGEN} from R package bigsnpr \cite[]{prive2017efficient}, one by reading the BGEN data at random according to imputation probabilities, and another one reading it as dosages (i.e.\ expected values according to imputation probabilities).
The first dataset is used as what could be the real genotype calls and the second dataset as what would be its imputed version; this technique was used in \cite{prive2019making}.
